---
title: "Untitled"
output: html_document
---

# Cellular architecture of human brain metastases
```{text, include=FALSE, eval=FALSE}

# # Cellular architecture of human brain metastases
# https://pubmed.ncbi.nlm.nih.gov/35063085

# # GEO dataset: GSE186344
# https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE186344

```


## Install if missing function
```{r}
installpackageifmissing = function( package ) {
 package = package[!(package %in% installed.packages()[,"Package"])]
 if(length(package)){ 
   message(cat('Installing ',package,'...'))
   install.packages(package,dependencies = TRUE,repos = "http://cran.us.r-project.org")
 }
 #lapply(package, require, character.only=T)
}
installpackageifmissing_bioc = function( package ) {
 package = package[!(package %in% installed.packages()[,"Package"])]
 if(length(package)){ 
   message(cat('Installing ',package,'...'))
   BiocManager::install(package,ask=FALSE)
 }
 #lapply(package, require, character.only=T)
}
```

# Install necessary packages
```{r}
# https://github.com/satijalab/seurat
install.packages("Seurat")

# For plotting data
installpackageifmissing("ggplot2")

installpackageifmissing("BiocManager")
# For calling cell types by SingleR and available reference cel types (celldex)
installpackageifmissing_bioc("SingleR")
installpackageifmissing_bioc("celldex")
installpackageifmissing_bioc("SingleCellExperiment")
installpackageifmissing_bioc("scuttle")

```

# Define directories and gather files 
```{r}

# File path to new directory for data
singleCellDir = path.expand('~/Downloads/singleCell')
dir.create(singleCellDir)

# Heterogeneous (malignant and stromal) melanoma tumor
GSMsample = "GSM5645890_Melan_3"
dotGov = "https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM5645nnn/GSM5645890/suppl"
tenXfilenames = c("matrix.mtx.gz", "features.tsv.gz", "barcodes.tsv.gz")

# Create variables with names of files to get and save as
web10xFileNames = file.path(dotGov, paste0(GSMsample, "_", tenXfilenames))
local10xFileNames = file.path(singleCellDir, GSMsample, tenXfilenames)

# Create directory for downloading data
dir.create(file.path(singleCellDir, GSMsample), recursive = T)

# Download 10x files with specific names
mapply(download.file, web10xFileNames, local10xFileNames)

# Download cell type call annotations
webAnnotationFileName = file.path(dotGov, paste0(GSMsample, "_Cell_Types_Annotations.csv.gz"))
localAnnotationFileName = file.path(singleCellDir, paste0(GSMsample, "_Cell_Types_Annotations.csv.gz"))
download.file(webAnnotationFileName, localAnnotationFileName)

# Check if files downloaded to folder (and named appropriately)
dir(file.path(singleCellDir, GSMsample))
dir(singleCellDir, pattern = "csv.gz")

# expect to see:
# "barcodes.tsv.gz" "features.tsv.gz" "matrix.mtx.gz"
# "GSM5645890_Melan_3_Cell_Types_Annotations.csv.gz"
```

# Read data into R
```{r}
library("Seurat")
library("ggplot2")

# Read annotation data from file
cellAnnotPaper = read.csv(dir(singleCellDir, pattern = "Annotations.csv.gz", full.names = T), row.names = 1)

# Read 10x data into R via Seurat
tempSampleData = Read10X(file.path(singleCellDir, GSMsample))

# Exclude genes expressed in less than 5 cells - described in paper
Melan_3 = CreateSeuratObject(counts = tempSampleData, project = "Melan_3", min.cells = 5)

# For space
rm(tempSampleData)

# Add the percent of genes counts that are from mitochondrial genes to the metadata
Melan_3[["percent.mt"]] <- PercentageFeatureSet(Melan_3, pattern = "^MT-")

# Plot some metadata features for each cell
# VlnPlot(Melan_3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(Melan_3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, log = T)

FeatureScatter(Melan_3, feature1 = "nFeature_RNA", feature2 = "nCount_RNA")
FeatureScatter(Melan_3, feature1 = "nFeature_RNA", feature2 = "percent.mt")
FeatureScatter(Melan_3, feature1 = "nCount_RNA", feature2 = "percent.mt")
```

## Add cell type calls from the paper to the Seurat object
```{r}
# view cell names from paper and Seurat object each
cellAnnotPaper[1:5, ,drop=F]
Melan_3@meta.data[1:5,]
rownames(Melan_3@meta.data)[1:5]
rownames(cellAnnotPaper)[1:5]

# Store original row names - helps prevent errors if the line is run again
cellAnnotPaper$orig_row_names = rownames(cellAnnotPaper)

# Capture cell barcodes
cellAnnotPaper$cellBarcode = gsub("^(1|G).*E13bm_(.*)$", "\\2", cellAnnotPaper$orig_row_names)

# Make names match 10x row names
cellAnnotPaper$CellBarcode10x = paste0(cellAnnotPaper$cellBarcode, "-1")

# Add paper's cell type calls to seurat object
rownames(cellAnnotPaper) = cellAnnotPaper$CellBarcode10x
meta.data = Melan_3@meta.data
meta.data$Cell_Type_paper = cellAnnotPaper[rownames(meta.data), "Cell_Type"]
meta.data$Cell_Type_paper[is.na(meta.data$Cell_Type_paper)] = "NA"
Melan_3 = AddMetaData(Melan_3, meta.data)
```

# Subset data by QC cutoffs (and cell type annotations)
```{r}
# Inspect cutoffs used in paper
FeatureScatter(Melan_3, feature1 = "nFeature_RNA", feature2 = "percent.mt") +
    ggplot2::geom_hline(yintercept = 10) + 
    geom_vline(xintercept = 500) + geom_vline(xintercept = 9000) + ylim(NA,100)

# Subset to only those with a cell type call from the paper
Melan_3_Pub = subset(Melan_3, subset = Cell_Type_paper != "NA")

FeatureScatter(Melan_3_Pub, feature1 = "nFeature_RNA", feature2 = "percent.mt") +
    ggplot2::geom_hline(yintercept = 10) +
    geom_vline(xintercept = 500) + geom_vline(xintercept = 9000)+ ylim(NA,100)
```

## Call cell type via SingleR and celldex reference
```{r}
# https://bioconductor.org/packages/release/bioc/vignettes/SingleR/inst/doc/SingleR.html
# https://bioconductor.org/packages/3.15/data/experiment/html/celldex.html

library("SingleR")
library("celldex")

# Get expression of reference cells
ref.BP <- BlueprintEncodeData() # melanocytes. skeletal muscle. melanoma and rhabdoid samples


# Prepare data for SingleR
tempSC = as.SingleCellExperiment(Melan_3_Pub)
# tempSC <- scuttle::logNormCounts(tempSC)

# Make cell type predictions
pred.Melan_3 = SingleR(test = tempSC, ref = ref.BP, assay.type.test=2, labels = ref.BP$label.main)

# Add SingleR cell type calls to seurat object's metadata
meta.data = Melan_3_Pub@meta.data
meta.data[ ,"cellCall"] = NA
meta.data[rownames(pred.Melan_3), "Cell_Type_SingleR"] = pred.Melan_3$labels
Melan_3_Pub = AddMetaData(Melan_3_Pub, meta.data)

```


## Perform seurat normalization and standard workflow
```{r}
# Before clustering, the number of counts and percentage of mitochondrial genes were regressed out (function vars.to.regress)

Melan_3_Pub <- NormalizeData(Melan_3_Pub, verbose = F)
Melan_3_Pub <- FindVariableFeatures(Melan_3_Pub, selection.method = "vst", nfeatures = 2000, verbose = F)
Melan_3_Pub <- ScaleData(Melan_3_Pub, vars.to.regress = c("nCount_RNA", "percent.mt"), verbose = F)
Melan_3_Pub <- RunPCA(Melan_3_Pub, features = VariableFeatures(Melan_3_Pub), verbose = F)
Melan_3_Pub <- RunUMAP(Melan_3_Pub, dims = 1:18, verbose = F)
```


## Plot cell types from publication and SingleR 
```{r}
DimPlot(Melan_3_Pub, reduction = "umap", group.by = "Cell_Type_paper")
DimPlot(Melan_3_Pub, reduction = "umap", group.by = "Cell_Type_paper", label = T, repel = T)

DimPlot(Melan_3_Pub, reduction = "umap", group.by = "Cell_Type_SingleR")
DimPlot(Melan_3_Pub, reduction = "umap", group.by = "Cell_Type_SingleR", label = T, repel = T)
```


## Plot common gene markers
```{r}
# CD45+ - immune cell marker
# CD4 - CD4+ T-cell
# CD8A - CD8+ T-cell
# MLANA - Melanoma

genesPlot = c("PTPRC", "CD4", "CD8A", "MLANA")

FeaturePlot(Melan_3_Pub, genesPlot)
```

## Compare numbers between cell type calls
```{r}

# Add Stromal malignant cell type calls
meta.data = Melan_3_Pub@meta.data

meta.data$malig_strom_paper = "stromal"
meta.data$malig_strom_paper[meta.data$Cell_Type_paper == "MTCs"] = "malignant"

meta.data$malig_strom_SingleR = "stromal"
meta.data$malig_strom_SingleR[meta.data$Cell_Type_SingleR == "Melanocytes"] = "malignant"

Melan_3_Pub = AddMetaData(Melan_3_Pub, meta.data)

# Check numbers of malignant and stroma between the paper's cell calls and SingleR 
table(Melan_3_Pub@meta.data[,c("malig_strom_SingleR", "malig_strom_paper")])


# Check specific cell type calls
table(Melan_3_Pub@meta.data[,c("Cell_Type_SingleR", "Cell_Type_paper")])

```


## Description of cell types called in the paper and the SingleR Blueprint reference
```{r, include=FALSE, eval=FALSE}
# All types in paper. All tumors:
# endothelial cells (EC)
# pericytes (PC)
# vascular smooth muscle cells (vSMCs)
# mesenchymal stromal cell-like cells (MSC-like)
# T cells (CD4+ CD8+ TReg T memoryT:CD8+:EM, T:CD4+, Tregs and T:CM)
# B cells (B)
# metastasis-associated macrophages (MAMs)
# dendritic cells (DCs)
# astrocytes

# All types in the Blueprint reference dataset
# sort(unique(ref.BP$label.main))
# "Adipocytes"        "Astrocytes"        "B-cells"           "CD4+ T-cells"      "CD8+ T-cells"      "Chondrocytes"      "DC"               
# "Endothelial cells" "Eosinophils"       "Epithelial cells"  "Erythrocytes"      "Fibroblasts"       "HSC"               "Keratinocytes"    
# "Macrophages"       "Melanocytes"       "Mesangial cells"   "Monocytes"         "Myocytes"          "Neurons"           "Neutrophils"      
# "NK cells"          "Pericytes"         "Skeletal muscle"   "Smooth muscle"  

# "DC" - Dendritic cells
# "HSC" - hematopoietic stem cells

# In Mel_3
# sort(unique(cellAnnotPaper$Cell_Type))
# "Astrocytes" "B-c1" "B-c2" "cDC2:CD1C+/CLEC10A+" "EC-1" "EC-2" "EC-3" "MAMs:APOE+" "MAMs:S100A8+" "MSC-like-c1" "MSC-like-c2" "MTCs" "PC-1" "PC-2" "PC-3" "T:CD4+:CM1" "T:CD4+:CM2" "T:CD8+:EM" "T:CM" "Treg" "vSMCs" 

# pericytes - "Pericytes are fibroblast-like cells with extensive cytoplasmic processes that wrap around endothelial cells in arterioles, capillaries, and venules" 
# - https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/pericyte

```


## Exercise - Plot additional gene markers from paper
```{r}
# # Stromal markers and curated gene signatures in supplemental file from paper
# https://www.sciencedirect.com/science/article/pii/S0092867421015658?via%3Dihub#mmc2

genesPlot = c("")
FeaturePlot(Melan_3_Pub, genesPlot)

```


## Exercise - generate tSNE plot
```{r}
Melan_3_Pub <- RunTSNE(Melan_3_Pub, dims = 1:18)
DimPlot(Melan_3_Pub, reduction = "tsne", group.by = "Cell_Type_SingleR", label = T, repel = T)

```


## Exercise - perform cell type calling using a different reference 
```{r, include=FALSE, eval=FALSE}
ref.HPCA <- HumanPrimaryCellAtlasData() # no melanocytes. Epithelial_cells:bronchial - for Lung_1


# Make cell type predictions


# Add SingleR cell type calls to seurat object's metadata



# Add Stromal malignant cell type calls



# Check numbers of malignant and stroma between the paper's cell calls and SingleR 



# Check specific cell type calls
```


## Remove cell type samples from blueprint reference and re-run SingleR - check if it improves malignant non-malignant cell type calling
```{r, include=FALSE, eval=FALSE}
# Remove cell types that you do not expect to appear in metastatic brain tumors

# Get names of cell types to keep
checkLabels = colData(ref.BP)
sort(unique(checkLabels$label.main))

# Exclude some cell types
labelsKeep = !checkLabels$label.main %in% c("Skeletal muscle")

# Reduce to cell types not including ILC and NKT
ref.BP.subset = subset(ref.BP, select = labelsKeep)


# Make cell type predictions


# Add SingleR cell type calls to seurat object's metadata



# Add Stromal malignant cell type calls



# Check numbers of malignant and stroma between the paper's cell calls and SingleR 



# Check specific cell type calls

```


## Exercise - broaden labels to non-immune and immune lineages
```{r, include=FALSE, eval=FALSE}
# Create new reference labels according to immune cell type lineages:
# melano
# non-immune
# Myeloid
# Lymphoid
# Googe search or see slide 7 of https://seer.cancer.gov/tools/heme/webinars/03_lineages1.pdf 

melano = c("Melanocytes")
nonimmune = c("Fibroblasts", "Keratinocytes", "Epithelial cells", "Skeletal muscle", "Smooth muscle") # not a complete list
lymphoidCells = c("B-cells", "CD4+ T-cells", "CD8+ T-cells", "NK cells")
myeloidCells = c()


# Get names of cell types to keep
checkLabels = colData(ref.BP)
sort(unique(checkLabels$label.main))

# Set new label names
checkLabels$label.broaden = ""
checkLabels[checkLabels$label.main %in% melano, "label.broaden"] = "melano"
checkLabels[checkLabels$label.main %in% nonimmune, "label.broaden"] = "non-immune"
checkLabels[checkLabels$label.main %in% lymphoidCells, "label.broaden"] = "lymphoid"
checkLabels[checkLabels$label.main %in% SOMETHING_HERE, "label.broaden"] = "SOMETHING_HERE"
colData(ref.BP) <- checkLabels

# Make cell type predictions



# Add SingleR cell type calls to seurat object's metadata



# Check numbers of malignant and stroma between the paper's cell calls and SingleR 


# Check specific cell type calls


# Add Stromal malignant cell type calls


# Check numbers of malignant and stroma between the paper's cell calls and SingleR 

```



# Check which do not match
```{r}

# # Check which do not match
# Melan_3_Pub@meta.data[Melan_3_Pub@meta.data$malig_strom_paper == "stromal" & Melan_3_Pub@meta.data$malig_stroma_SingleR == "malignant",]
# Melan_3_Pub@meta.data[Melan_3_Pub@meta.data$malig_strom_SingleR == "stromal" & Melan_3_Pub@meta.data$malig_strom_paper == "malignant",]


```

