---
title: "Data Basics Assignment"
author: "Tony Kabilan Okeke"
date: "January 10, 2023"
---

In this assignment you are going to analyze this [Immunotherapy dataset](http://archive.ics.uci.edu/ml/datasets/Immunotherapy+Dataset). The study that has produced the dataset is described [here](https://www.ncbi.nlm.nih.gov/pubmed/28086200). The dataset contains only the immunotherapy subset of patients. Review the paper for description of the attributes.

- Create a `hwdatabasics.mlx`, `hwdatabasics.ipynb` or `hwdatabasics.rmd` file
  to perform the following tasks.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set the working directory to the location of the data file
setwd("/home/muaddib/sietch/courses/bmes547/1_data/homework")
```

```{r packages message=FALSE, warning=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse)
```

```{r data}
# Download and read the data
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00428/Immunotherapy.xlsx"
if (!file.exists("immunotherapy.xlsx")) download.file(url, "immunotherapy.xlsx")
data <- readxl::read_excel("immunotherapy.xlsx")
```

- For each attribute, find its correlation with the result of treatment. Exclude 
  result of treatment as an attribute so you do not find its correlation with 
  itself. Note that correlation is not the best measurement for categroical 
  features, especially if there are more than 2 categories. But, for this 
  assignment, we'll ignore this caveat and calculate correlation anyway. You do 
  not need to convert categorical features for this assignment, just use the 
  numbers as they are.

```{r correlation}
# Calculate correlation with the result of treatment
corr <- cor(data, use = "complete.obs")[1:7, 8]

# Store the correlation values in a data frame
results <- tibble(name=names(corr), corr)
```

- For each attribute, find if there is a statistically significant difference in 
  that attribute, between successful and failed treatments. Use 
  *two-sample t-test* for the comparison and find the p-value. Report the names 
  of attributes that you found to be statistically significantly different. You 
  need to decide your own p-value threshold.

```{r t-test}
# Test for statistically significant difference between groups
# Assume that the data is normally distributed, and variance is equal
pvals <- data %>%
    pivot_longer(-Result_of_Treatment) %>%
    group_by(name) %>%
    rstatix::t_test(
        value ~ Result_of_Treatment, paired = FALSE, var.equal = TRUE
    ) %>%
    select(name, p)

# Join to the results data frame
results <- results %>%
    left_join(pvals, by = "name")

# Report the statistically significant attributes
p_thr <- 0.05
sig <- results %>%
    filter(p < p_thr) %>%
    .$name
paste("The following attributes are statistically significant:", 
      paste(sig, collapse = ", "))
```

- Report the correlation coefficients and p-values of **all of** the attributes 
  in a tabular format (First column containing variable names, second column 
  containing correlation values, and third column containing pvalues). The 
  reported table must be sorted in increasing order by p-values.

```{r table}
# Display the results in a table
results %>%
    arrange(p) %>%
    knitr::kable()
```

- For the attribute with the most significant p-value, create a box plot showing 
  a summary of the values of that attribute, for successfule and failed 
  treatments.

```{r boxplot}
# boxplot of the attribute with the most significant p-value
top <- results %>%
    arrange(p) %>%
    .$name %>%
    .[1]
data$Result_of_Treatment <- factor(data$Result_of_Treatment)

data %>%
    ggplot(
        aes(x=Result_of_Treatment, y=!!sym(top), fill=Result_of_Treatment)
    ) +
        geom_boxplot()
```
