---
format:
  pdf:
    documentclass: scrartcl
    fig-width: 4
    fig-height: 2.5
    fontfamily: FiraSans
    fontsize: 11pt
    geometry: 
      - top=0.7in
      - bottom=1in
      - left=0.5in
      - right=0.5in
      - heightrounded
    highlight-style: github
    include-in-header: 
      text: |
        \usepackage[shortlabels]{enumitem}
        \usepackage{fancyhdr}
        \pagestyle{fancy}
        \lhead{Tony Kabilan Okeke}
        \chead{Assignment 8}
        \rhead{BMES 315}
    mainfont: Lato
    papersize: letter
---

<!-- \vspace{-1.45in} -->

```{r}
#| echo: false
#| warning: false
#| message: false

# Load packages
pacman::p_load(tidyverse)

# Custom functions
plot.lm <- function(data, x, y, xlab, ylab, main) {
    # Plot the linear model
    ggplot(data, aes_string(x = x, y = y)) +
        geom_point(size = 2) +
        geom_smooth(method = "lm", formula = y~x, color = "red",
                    fill = "#db8e88", se=FALSE) +
        ggpubr::stat_regline_equation(aes(
            label = paste(..eq.label.., ..adj.rr.label.., sep = "~~~")
        )) +
        labs(title = main, x = xlab, y = ylab) +
        ggpubr::theme_pubr() +
        theme(
            plot.title=element_text(hjust=0.5, size=13),
            axis.title=element_text(size=12),
            axis.text=element_text(size=10),
            text=element_text(family="Helvetica"),
        )
}

pred <- function(fit, cap, ...) {
    # Predict the fitted values
    x <- predict(fit, newdata=data.frame(...))
    paste("For a student with ", cap, ", GPA = ", round(x, 2), sep = "")
}

# Load data set
df <- tibble(read.csv("hw8-data.txt", sep='\t'))
```


# Question 1
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

Data was collected after the end of an introductory statistics course in order 
to investigate the relationship between certain variables, such as study time
per week, and overall GPA. The table presented below shows the data for 8
females in the class on these variables.

\begin{enumerate}[(a)]
\item {Create a scatter plot between study time and college GPA}
\item {Analyze the data and report the equation and slope of the regression line}
\item {Interpret the results – what information should be given to students
       concerning the impact of study time on academic performance?}
\item {Find the predicted GPA for a student who spends 25 hours a week studying}
\end{enumerate}

\vspace{-5mm}
```{r}
#| echo: false

# Show the data
knitr::kable(df, "pipe", align="c")
```

:::

```{r}
#| echo: true

# Fit the data to a linear model
fit <- lm(gpa ~ study.time, data=df)
summary(fit)
```

::: {.callout-tip appearance="minimal"}

Based on the results of the analysis, there appears to be an increase in the
overall GPA of students as their weekly study time increases 
$\left(R_{adj} = 0.60, p=0.015\right)$; specifically, for every additional hour
of study time, there is an increase in GPA of 0.04.  

The students should be informed that studying for a greater amount of time per
week tends to increase their overall GPA.

:::

```{r}
# Predict the GPA for a student who spends 25 hours a week studying
pred(fit, "a study time of 25 hours", study.time = 25)
```

```{r fig.align='center'}
#| echo: false

plot.lm(df, "study.time", "gpa", "Study Time (hours/week)", "GPA", 
        "Study Time vs. GPA")
```

# Question 2
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

Repeat a-c from above with number of lectures missed as the predictive, x,
variable. For part d, find the predicted GPA for a student who missed 12
lectures.

:::

```{r}
#| echo: true

# Fit the data to a linear model
fit <- lm(gpa ~ missed.lectures, data=df)
summary(fit)
```

::: {.callout-tip appearance="minimal"}

Based on the results of the analysis, there appears to be a decrease in 
overall GPA of students as the number of lectures they miss increases
$\left(R_{adj} = 0.77, p=0.002\right)$; specifically, for every lecture missed,
there students' GPA decreases by 0.082.  

The students should be informed that missing lectures tends to decrease their
overall GPA.

:::

```{r}
# Predict the GPA for a student who missed 12 lectures
pred(fit, "12 missed lectures", missed.lectures = 12)
```

```{r fig.align='center'}
#| echo: false

plot.lm(df, "missed.lectures", "gpa", "Missed Lectures", "GPA", 
        "Missed Lectures vs. GPA")
```


# Question 3
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

Repeat a-c from above with average daily sleep time as the predictive, x,
variable. For part d, find the predicted GPA for a student who averaged 7.8
hours of sleep.

:::

```{r}
#| echo: true

# Fit the data to a linear model
fit <- lm(gpa ~ sleep.time, data=df)
summary(fit)
```

::: {.callout-tip appearance="minimal"}

Based on the results, there appears to be an increase in overall GPA with an
increase in the number of hours spent sleeping 
$\left(R_{adj} = 0.97, p=3.3\cdot10^{-6}\right)$; specifically, the closer a
student's sleep time is to 8 hours, the higher their GPA tends to be.
Furthermore, average daily sleep time was found to be responsible for 97% of
the variation in GPA.  

The students should be informed that students who sleep for closer to 8 hours
per day tend to have higher GPAs.

:::

```{r}
# Predict the GPA for a student who missed 12 lectures
pred(fit, "an average of 7.8 hours of sleep each day", sleep.time = 7.8)
```

```{r fig.align='center'}
#| echo: false

plot.lm(df, "sleep.time", "gpa", "Sleep Time (hours/day)", "GPA", 
        "Sleep Time vs. GPA")
```


# Question 4
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

Consider the study as described. Would this receive IRB approval? If not, why
not and how could the study be changed so as to make the study more likely to
be approved?

:::

::: {.callout-tip appearance="minimal"}

I think it is unlikely that the study would receive IRB approval as it is
currently designed. For one, the study only includes 8 female students who
were all enrolled in an introductory statistics course. Also, The data collected
was from a convenience samlple, which is not representative of the population of
students at the university. 
In order to improve the study, I would suggest that the researchers collect
a stratified random sample of students from a variety of different majors, 
including multiple genders. In addition, the researchers would have to
more clearly define how study time and average daily sleep time are calculated
for each student.

:::


# Question 5
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

The above study was observational. Choose one of the factors and create an
experimental study. Would your design be likely to receive IRB approval? What
might some of the confounding factors be in your study design and how might
you control for them?

:::

::: {.callout-tip appearance="minimal"}

**Experimental Design**  

- **Sampling:** Take a stratified random sample of 100 students in their freshman
  year (all genders included) from the College of Engineering at Drexel
  University who are currently enrolled in classes. This should be done at the
  start of an academic quarter. Students should be stratified by gender.
    - The selected students should all be enrolled in an introductory statistics
      course which meets twice a week for 1 hour each.
    - The students should be randomly assigned to five groups of 20 students each.
        - Group 1: Students will be instructed to attend 100% of the lectures
        throughout the quarter.
        - Group 2: Students will be instructed to attend 75% of the lectures
        throughout the quarter.
        - Group 3: Students will be instructed to attend 50% of the lectures
        throughout the quarter.
        - Group 4: Students will be instructed to attend 25% of the lectures
        throughout the quarter.
        - Group 5: Students will be instructed not to attend any lectures
        throughout the quarter.
        - Students will be allowed to select which lectures they attend.
    - *Sample Size*, $n = 20 ~\forall~ \text{groups}$ 

- **Measurements:** The students performance in the course will be measured
  based on their performance on the midterm and final exams. The midterm and
  final will be weighted equally in the final grade. The midterm will cover
  material from weeks 1-5 and the final will cover material from weeks 6-10.
  Students overall grade will be calculated as the average of their midterm and
  final exam scores.

- **Analysis:** The students' performance in the course will be compared across
  the five groups. The students' performance will be measured by their overall
  grade in the course. The data will be analyzed using a two-way ANOVA with
  the factors being number of lectures attendend and gender. If the ANOVA
  indicates that there is a significant difference in the students' performance
  across the five groups, then a post-hoc test will be performed to determine
  which groups are significantly different from each other.

- **Note:** This study will narrow the scope of the original study by only
  focusing on the effect of lecture attendance on the students' performance in
  a single course, instead of on their overall GPA.


**IRB Approval**

This study is unlikely to be approved by an IRB because there may not be an
ethical way to fully control the sleep time of the students. Whereas study time
could theoretically be controlled by the researchers, it would be much harder to
require students to sleep for a certain amount of time each day throughout an 
entire quarter; depending on the nature of their courseload and extracurricular
activities, students are likely to have varying amounts of sleep. Furthermore, 
it would likely be an invasion of the students' privacy to implement any strict
monitoring system for their sleep schedules throughout the 10 weeks of the term. 
Additionally, since the study is only being conducted on students in the College
of Engineering taking an introductory statistics course, the results of the
study may not be generalizable to the entire student population at Drexel
University. Finally, since the study could have a negative impact on the
students' GPA which could have consequences on their emotional and mental
health, and may impact their job prospects since GPA is often used as a
screening tool by employers.


**Confounding Factors**

- Students may have varying levels of prior experience with the material covered
  in the course

- The amount of time spent studying for the course may vary across the groups
  which would likely impact their performance in the course. To control for
  this, students across all groups should be required to study for an average
  of 12 hours per week. For the purposes of this study, "study time" will be
  defined as the number of hours spent working on course material outside of
  class time.

- The amount of sleep time (as shown in our analyses above) has a significant
  impact on students' overall GPA. To control for this, students across all
  groups should be required to sleep for an average of 8 hours per day.

- There students' performance could also be impacted by variations in course
  instructors, teaching assistants, and the lecture material used. To control
  for this, all students in the study should be taught the same material by the
  same instructor and teaching assistants throughout the quarter.

:::


# Question 6
\vspace{-5mm}

::: {.callout-note appearance="minimal"}

Run the data table as a multiple regression using all three predictive variables
and GPA as the variable being predicted. Do the conclusions change? How?

:::

```{r}
#| echo: true

# Fit the data to a linear model
fit <- lm(gpa ~ sleep.time + study.time + missed.lectures, data=df)
summary(fit)
```

::: {.callout-tip appearance="minimal"}

When we look at a multiple regression model, only daily sleep time was found
to be a significant predictor of GPA. This is because the other variables
were likely confounding variables and the correlation between them and GPA
was spurious.

:::
